---
output:
  pdf_document: default
  html_document: default
---

## Data 605-Assignment 4 ##

**Hazal Gunduz**

**With the attached data file, build and visualize  eigenimagery that accounts for 80% of the variability.  Provide full R code and discussion.**

**Loading Libraries**

```{r}
library(foreach)
library(jpeg)
library(EBImage)
library(OpenImageR)
```

*We have 17 shoes images in the .jpg folder.*

```{r}
files<-list.files("/Users/otheraccount/Downloads/jpg/", pattern="\\.jpg")[1:17] 
```

**Shoes images function**

```{r}
height = 1200; 
width = 2500; 
scale = 20
plot_jpeg = function(path, add=FALSE) { 
jpg = readJPEG(path, native=T) 
res = dim(jpg)[2:1] 
  if (!add) 
  plot(1,1,xlim=c(1,res[1]), ylim=c(1, res[2]), asp=1, type='n', xaxs='i', yaxs='i', xaxt='n', yaxt='n', xlab='', ylab='', bty='n')
  rasterImage(jpg, 1, 1, res[1], res[2])
}
```

*Loading the data into an array and resizing function scales.*

```{r}
images<-array(rep(0, length(files)*height/scale*width/scale*3), dim=c(length(files), height/scale, width/scale, 3))

for (i in 1:17) {
temp<-resize(readJPEG(paste0("/Users/otheraccount/Downloads/jpg/", files[i])), height/scale, width/scale)
images[i,,,]=array(temp, dim=c(1, height/scale, width/scale, 3))
}
dim(images)
```

**Vectorizing**

```{r}
flat<-matrix(0, 17, prod(dim(images))) 

for (i in 1:17) {
newimages<-readJPEG(paste0("/Users/otheraccount/Downloads/jpg/", files[i]))
r=as.vector(images[i,,,1]); g=as.vector(images[i,,,2]); b=as.vector(images[i,,,3])
flat[i,] <- t(c(r, g, b))
}

shoes<-as.data.frame(t(flat))
```

**Visualize the actual plots**

*Let’s look at the shoes, using the 'plot_jpeg' function to see the images.*

```{r}
par(mfrow=c(6,3))
par(mai=c(.03,.03,.03,.03))

for (i in 1:17) { 
plot_jpeg(writeJPEG(images[i,,,]))
}
```

**Structure of Eigencomponents**

*We need to scale and center data with using the scale function.*

```{r}
scaled<-scale(shoes, center=TRUE, scale=TRUE)
mean.shoe<-attr(scaled, "scaled:center")
std.shoe<-attr(scaled, "scaled:scale")
dim(scaled)
```

**Calculate Covariance (Correlation)**

*We are using the Correlation matrix. And correlation matrix will be used to determine the eigencomponents.*

```{r}
Sigma_<-cor(scaled)
dim(Sigma_)
```

**Get the Eigencomponents**

```{r}
myeigen<-eigen(Sigma_)
myeigen
```

```{r}
myeigen<-eigen(Sigma_)
cumsum(myeigen$values) / sum(myeigen$values)
```

*We see that 80% variability at the position of 2, so we find the eigenshoes at [, 2].*

**Eigenshoes**

```{r}
scaling<-diag(myeigen$values[1:5]^(-1/2)) / (sqrt(nrow(scaled)-1))
eigenshoes<-scaled%*%myeigen$vectors[,1:5]%*%scaling
par(mfrow=c(2,3))
imageShow(array(eigenshoes[,2], c(60, 125, 3)))
```

*Transform of the images.*

```{r}
height=1200
width=2500
scale=20
newimages=images
dim(newimages)=c(length(files), height*width*3/scale^2)
pca<-princomp(t(as.matrix(newimages)), scores=TRUE, cor=TRUE)
```

**Eigenshoes**

```{r}
pca2<-t(pca$scores)
dim(pca2)=c(length(files), height/scale, width/scale, 3)
par(mfrow=c(5,5))
par(mai=c(.01,.01,.01,.01))
for (i in 1:17) {
plot_jpeg(writeJPEG(pca2[i,,,]))
}
```

```{r}
a<-round(pca$sdev[1:17]^2 / sum(pca$sdev^2), 3)
cumsum(a)
```

**New Data Set**

We will write the new data set 80% of the variability for all images.

```{r}
x <- t(t(eigenshoes)%*%scaled)
```


Rpubs ->  https://rpubs.com/gunduzhazal/1005162

